Road Segment Prioritization for Bicycle Infrastructure

### 2021-01-09 README.Rmd
edit section 7 + 8

All of these network-level methodologies are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors. This is important since different ethical principles constitute different problem formulations and targets. Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles. The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport]. @nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals. @lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area. @pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning. They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility. This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling. **Write some more here?**

@@

After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments. This cumulative demand (referred to as *flow*) is then used as a basis for determining where best to invest in segregated cycling infrastructure. In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.

For this purpose, two algorithms are proposed. Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism. In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded. Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.

@@

While segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum. @wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%. However, cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@pucher2010infrastructure]. These cities do not just focus on infrastructure, but on general policies as well as restricting car use. Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making]. While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.

@@

This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study. 
The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can be a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].

>>>>

All of these network-level methodologies are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors. This is important since different ethical principles constitute different problem formulations and targets. Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles. The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport]. @nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals. @lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area. @pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning. They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility. This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling.

@@

After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments. This cumulative demand (referred to as *flow*) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure. In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well-connected routes.

A range of algorithms could be used for prioritisation. Because policy priorities vary, we present two algorithms. Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism. In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded. Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.

@@

While segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum. @wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%. International research shows that cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure]. These cities do not just focus on infrastructure, but on general policies as well as restricting car use. Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making]. While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.

@@

This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study. 
The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can provide an evidence-based starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. Given that cycle network development is seen as the “most essential activity entailed in the design of cycle-friendly infrastructure” [@parkin2018designing], we believe that the methods have great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term. A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and origin-destination data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. For instructions on reproducing the results shown in this paper, see the **README** file in the R folder of the .zip file that accompanies this paper (to be released as open source software when the paper is published).].

<!-- For after publishing

^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]. -->
